EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-28 03_edipack2.tex
Updates from Overleaf
Updated KMflake figure

The interaction part of the impurity Hamiltonian, 
 $\hat{H}^{\mathrm{int}}$ can in principle contain any 
set of two-body operators: 
$$
\hat{H}^\mathrm{int}_\mathrm{imp} =  \sum_{ijkl} U_{ijkl}c^{\dagger}_{i}c^{\dagger}_{j}c_{l}c_{k}. 
$$
However, we typically 
adopt a generic formulation of the local multi-orbital Hubbard-Kanamori interaction~\cite{Georges2013ACMP}:
\begin{equation}\label{Hint}
  \begin{split}
    \hat{H}^\mathrm{int}_\mathrm{imp} &=U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}\\
    &{\phantom =}- J_X\sum_{\a\neq
      \b}d^{+}_{\a\uparrow}d_{\a\downarrow}d^{+}_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a
      \neq
      \b}d^{+}_{\a\uparrow}d^{+}_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow}\\
\end{split}
\end{equation}
where we introduced the occupation number operator   $n_{\alpha\sigma}=d^{+}_{\alpha\sigma}d_{\alpha\sigma}$.
The first three terms represent the density-density part of the
interaction, where $U$ is the local intra-orbital Coulomb repulsion,
$U'$ the inter-orbital one and $J$ the Hund's coupling~\cite{Georges2013ACMP}.  
The last two terms are, respectively, the spin-exchange ($J_X$) and
the pair-hopping ($J_P$).
In the three-orbital case $N_\a=3$ a fully symmetric $SU(3)_\mathrm{orbital}\otimes SU(2)_\mathrm{spin}\otimes
U(1)_\mathrm{charge}$ form of the interaction is obtained by setting $U'=U-2J$ and
$J_X=J_P=J$~\cite{Georges2013ACMP}. Different choices, preserving part
of the combined symmetry group, can be made for other number of orbitals~\cite{Georges2013ACMP}.

@@

\subsubsection{Electron-Phonon coupling}
Finally, the electron-phonon part of the quantum impurity problems is
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_\mathrm{ph}&=\sum_q \omega_{0q} b_q^+b_q + \sum_q \sqrt{2}A_{q}(b^\dagger_q + b_q)  \\
    % 
    \hat{H}_\mathrm{e-ph} &= \sum_q\sum_{\a\b\sigma} g_{\a\b} d^+_{\a\sigma}d_{\b\sigma}(b_q+b_q^+)
\end{split}
\end{equation}
where $q=1,\dots,N_q$ indexes the number of local phonons, and $b_q$
($b_q^+$) are the destruction (creation) operators for a phonon with
frequency $\omega_{0q}$.
The matrix  $g_{\a\b}$ describe the electron-phonon coupling. 
Although feasible, dealing with more than one phonon mode becomes
quickly computationally very demanding, thus in the rest of the this
work we shall consider $N_q=1$.
The phonon's frequency is passed to the \NAME program via the variable { \tt W0\_PH } to be read from the input file. The displacement field acting on the phonon is passed through the variable {\tt A\_PH} from the input file. The electron-phonon coupling matrix $g_{\alpha \beta}$ instead can be passed in two different ways. If the matrix is diagonal ($g_{\alpha \beta} = g_\alpha \delta_{\alpha \beta}$), from the input file one can simply write in order the values $g_\alpha$ in the variable {\tt G\_PH }. For any general shape one can pass to the input file a {\tt GPHfile } variable containing the name of the file storing the $g_{\alpha \beta}$ array in Fortran style.
If {\tt GPHfile=NONE} the matrix $g_{\alpha \beta}$ is read from the variable {\tt G\_PH}, else it is read from the {\tt GPHfile}.

\subsubsection{System setup}
We consider a quantum impurity system composed by a single multi-orbital impurity, a bath discretized into a number $N_\mathrm{bath}$ of {\it elements} possibly endowed with an internal structure, and a number of available phonons. 
The total size of the system is determined by three contributions: i) the number of impurity orbitals $N_\a$, ii) the total count of electronic bath levels $N_\mathrm{b}$, and iii) by the number of phonons $N_\mathrm{ph}$. 
The latter is fixed to set a cut-off to the unbounded dimensions of the local phonons Hilbert space.
The number of bath degrees of freedom $N_\mathrm{b}$ is a
function of the bath topology and of $N_\mathrm{bath}$, which is determined internally by \NAME.
In the simplest case, each bath
element corresponds to an independent electronic level coupled to the
impurity, thus $N_\mathrm{b}\equiv N_\mathrm{bath}$.
The number of electronic levels in the quantum impurity systems is indicated with $N_\mathrm{s}$.

@@

% \paragraph{{\bf Remark}} As extensively discussed in
% \onlinecite{Amaricci2022}, the inclusion of local phonons with
% truncated dimensions ultimately amounts to a sequential application of
% the procedures defined for the electronic part, i.e. one per phonon
% mode. This includes a linear scaling of the dimensions with $N_\mathrm{ph}$
% and of course largely limit the available degrees of freedom to
% describe electronic states. In the rest of the paper we focus
% specifically on the electronic part of the quantum impurity problem,
% referring to the phonons in specific sections when their presence
% introduces non-trivial modifications.

\subsection{The Fock basis states}\label{sSecBasis}
The Fock space of the quantum impurity problem is defined as
$\FF=\FF_\mathrm{e}\otimes \FF_\mathrm{ph}$, where: $$
\FF_\mathrm{e}=\bigoplus_{n=0}^{N_\mathrm{s}}
S_-\HH_\mathrm{e}^{\otimes n}
$$ 
is the electronic Fock space built 
from the local electronic Hilbert space $\HH_\mathrm{e}$, and 
$$
\FF_\mathrm{ph}=\bigoplus_{n=0}^{N_q}S_+\HH_\mathrm{ph}^{\otimes n}
$$
is the
phonon Fock space, with local phonon Hilbert space
$\HH_\mathrm{ph}=\{\ket{0},\ket{1},\dots,\ket{N_\mathrm{ph}}\}$. 
Here ($S_-$) $S_+$ is the (anti-)symmetrization operator.  
%
The total dimension of the Fock space is
$$
D=4^{N_\mathrm{s}}\cdot (N_\mathrm{ph}+1)=D_\mathrm{e}\cdot
D_\mathrm{ph},
$$ 
highlighting the exponential
growth with the number of electronic levels.

@@

The electronic part of the Fock state $\ket{\vec{n}}$ is represented
as a binary string of length $2N_\mathrm{s}$. Thus, any such state can be encoded
in a computer using a sequence of $2N_\mathrm{s}$ bits, or equivalently, as an integer $I=0,\dots 2^{2N_\mathrm{s}}-1$ such that $\ket{\vec{n}}=\ket{I}$.

\fixme{Destruction and creation operators, $c_{a\sigma}$ and $c^+_{a\sigma}$ 
respectively, act on the Fock space as}:
\begin{align*}
  c_{a\sigma}\ket{\vec{n}} &=
    \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!-\!1,\dots}
      &\text{if $n_{a\sigma}\!=\! 1$}\\
      0 &\text{otherwise}
    \end{cases};\\
    c^{+}_{a\sigma}\ket{\vec{n}} &=
     \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!+\!1,\dots}
      & \text{if $n_{a\sigma}\!=\! 0$}\\
      0 & \text{otherwise}
    \end{cases}    
\end{align*}
where $\#_{a\sigma}=\sum_{b\sigma'<a\sigma} n_{b\sigma'}$ accounts for the fermionic sign imposed by the Pauli principle.

The implementation of Fock states and operators can be found in 
{\tt ED\_AUX\_FUNX}. In particular, in this module we define the bitwise action 
of fermionic creation and annihilation operators as functions 
{\tt CDG} and {\tt C}, used throughout the code. Additionally, 
we provide the function {\tt bdecomp} for reconstructing the Fock 
state bit sequence from the integer representation.

\begin{lstlisting}[style=fstyle,numbers=none]
  subroutine C(pos,in,out,fsgn) !$c_\mathrm{pos}|\mathrm{in}\rangle=|\mathrm{out}\rangle$
    integer,intent(in)    :: pos  !Position to be set to 1
    integer,intent(in)    :: in   !Input index
    integer,intent(inout) :: out  !Output index
    real(8),intent(inout) :: fsgn !Fermionic sign   
    integer               :: l
    if(.not.btest(in,pos-1))stop "C error: C_i|...0_i...>"
    fsgn=1d0
    do l=1,pos-1
       if(btest(in,l-1))fsgn=-fsgn
    enddo
    out = ibclr(in,pos-1)
  end subroutine C
\end{lstlisting}

\begin{lstlisting}[style=fstyle,numbers=none]
  subroutine CDG(pos,in,out,fsgn) !$c^\dagger_\mathrm{pos}|\mathrm{in}\rangle=|\mathrm{out}\rangle$
    integer,intent(in)    :: pos  !Position to be set to 1
    integer,intent(in)    :: in   !Input index
    integer,intent(inout) :: out  !Output index
    real(8),intent(inout) :: fsgn !Fermionic sign   
    integer               :: l
    if(btest(in,pos-1))stop "C^+ error: C^+_i|...1_i...>"
    fsgn=1d0
    do l=1,pos-1
       if(btest(in,l-1))fsgn=-fsgn
    enddo
    out = ibset(in,pos-1)
  end subroutine CDG
\end{lstlisting}

@@

\subsection{Conserved quantum numbers}\label{sSecQNs}
To mitigate the exponential scaling of the \fixme{Fock space dimension}, it 
is essential to exploit suitable symmetries. Specifically, if the 
Hamiltonian $H$ commutes with a set of operators ${\cal \vec{Q}}=[{\cal Q}_1,{\cal Q}_2,\dots,{\cal Q}_M]$ such that 
$[H, {\cal Q}_i] = 0$, the Fock space can be decomposed into smaller, 
disjoint symmetry sectors labeled by quantum numbers $\vec{Q}$.

In the context of quantum impurity problems, two commonly used 
symmetries are: i) conservation of the total charge $N$ and ii) 
conservation of the total magnetization $S_z$. Although the total 
spin operator $S^2$ may also be conserved, its implementation is 
computationally challenging and typically provides only marginal 
gains, making it less practical in many cases.

In \NAME, three different symmetry configurations are available, 
controlled by the input variable {\tt ed\_mode} = {\bf normal}, 
{\bf superc}, or {\bf nonsu2}:

\begin{itemize}
\item{} The {\bf normal} case conserves both the total occupation $N$
and the total magnetization $S_z$, or equivalently, the total number
of electrons with spin up $N_\up$ and down $N_\dw$.
Optionally, the symmetry
can be extended to act indipendently to each orbital and spin
component, i.e.~$\vec{N_\s}=[N^1_\s,\dots,N^{N_\a}_\s]$. This
specific case is discussed extensively
Ref.~\cite{Amaricci2022} so we will not be further cover it here recalling that any consideration concerning the {\tt normal} case extends directly to this case as well.

@@

\item{} The {\bf nonsu2} case conserves the total
charge $N$ while allowing spin symmetry group to be not conserved.
This scenario catpures effects such as local spin-orbit coupling
$\vec{L}\cdot\vec{S}$~\cite{something}, in-plane spin ordering~\cite{BellomiaKMH} or
in-plane spin-triplet exciton condensation~\cite{Amaricci2023_excitons,Blason}.  
\end{itemize}

From a computational perspective, constructing a symmetry
sector amounts to defining a injective map
$\MM:\SS_{\vec{Q}}\rightarrow \FF$ which relates the states $\ket{i}$
within a given symmetry sector $\SS_{\vec{Q}}$ to the states $\ket{I}$ in the
full Fock space.
This map is typically implemented as a rank-1
integer array whose size corresponds to the dimension of the sector $D_{\SS_{\vec{Q}}}$.

The {\bf normal} case deserves a special note. Since $N_\up$ and
$N_\dw$ are conserved independently, the local Hilbert
space and the electronic Fock space can be factorized as
$\HH_e=\HH_{e\up}\otimes\HH_{e\dw}$ and $\FF_\mathrm{e} = \FF_{e\up}\otimes \FF_{e\dw}$, respectively.  
Consequently, each Fock state can be written as a product
$\ket{\vec{n}_\up} \ket{\vec{n}_\dw}$. This factorization splits the
symmetry sector as $\SS_{\vec{Q}} = \SS_{N_\up}\otimes\SS_{N_\dw}$,
and finally the sector map can be expressed as the product $\MM = \MM_\up
\otimes \MM_\dw$.
%
Each state $\ket{i}=\ket{i_\up} \ket{i_\dw}$ in this factorized basis
is labeled by two integers $[i_\up,i_\dw]$, with 
$i_\sigma=1,\dots,D_{\SS_\sigma}$ such that $i=i_\up + i_\dw
D_{\SS_\dw}$.
The maps $\MM_\sigma$ then connect these basis states to Fock states
$\ket{I}=\ket{I_\up}\ket{I_\dw}$, labeled by two integers
$[I_\up,I_\dw]$ as $I=I_\up +   I_\dw 2^{N_\mathrm{s}}$.
For a more
detailed discussion on the structure of the Fock basis in this case, see Ref.~\cite{Amaricci2022}.

The presence of a symmetry reduces the Hamiltonian matrix to
a block diagonal form, where each block, labeled by $\vec{Q}$, has dimension
$D_{\SS_{\vec{Q}}}$. The sector Hamiltonian matrix $H_\SS$ is represented in the
basis $\ket{i}\in\SS_{\vec{Q}}$ as a sparse matrix.
In the {\bf normal} case this block structure is particularly
symmetric due to the factorized nature of the sectors as discussed
above, see also Ref.~\cite{Amaricci2022}.
The analysis of the spectrum is then reduced to inspecting the
Hamiltonian within each symmetry sector, or to a subset thereof if a
additional constraints are present.
A summary of the sector properties is provided in
Table~(\ref{TabSector}). 
\begin{table}%[ht]
\begin{center}
\begin{tabularx}{\linewidth}{ |X|X|X| } 
 \hline
  {\tt ed\_mode} & {\it Quantum Numbers} & {\it Sector Dimension} \\
  \hline
  {\bf normal} & $[N,S_z]\equiv[N_\up,N_\dw]$ &
                                                $\binom{N_\mathrm{s}}{N_\up}\binom{N_\mathrm{s}}{N_\dw}$
  \\
  \hline
  {\bf superc} & $S_z\equiv N_\up-N_\dw$ &  $\sum_m 2^{N_\mathrm{s}-S_z-2m}\binom{N_\mathrm{s}}{N_\mathrm{s}-S_z-2m}\binom{S_z+2m}{m}$
  \\
  \hline
  {\bf nonsu2} & $N \equiv N_\up+N_\dw$ & $\binom{2N_\mathrm{s}}{N}$ \\ 
 \hline
  \label{TabSector}
\end{tabularx}
\end{center}
\caption{
  {\bf Quantum Numbers Table}.
A table summarizing the possible values of the input variable {\tt
  ed\_mode} selecting the symmetries of the problem. To each value
(first column) correspond different quantum numbers (second column).
The third column report the dimension of the electronic symmetry sector $\SS_{\vec{Q}}$ for a
given value of the quantum numbers $\vec{Q}$.}
\end{table}

@@

In \NAME the implementation of symmetry sectors is managed through the
{\tt sector} object, which is defined in {\tt
  ED\_VARS\_GLOBAL}. This object contains all the relevant
information for defining the symmetry, including the sector
dimensions, quantum numbers and the map $\MM$. The constructor
(destructor) for this object is defined in the {\tt
  ED\_SECTORS} module via the function {\tt build\_sector} ({\tt
  delete\_sector}).
These functions use different algorithms depending on the nature 
of the quantum numbers $\vec{Q}$, i.e. the value of {\tt ed\_mode}. The following  code snippet summarize the various available implementations:

@@

\begin{tabular}{c|c|c}\label{list1}
    {\bf normal} & {\bf superc} & {\bf nonsu2}\\
    \hline
\stackinset{l}{}{t}{}{}{\listingA} &
\stackinset{l}{}{t}{}{}{\listingB} &
\stackinset{l}{}{t}{}{}{\listingC} \\
\end{tabular}
\vspace{2mm}

@@

\subsection{Interaction setup}\label{sSecIntSetup}
The \NAME code offers two distinct methods to setup the interaction terms $\hat{H}^\mathrm{int}$ in \equ{Himp}, which are controlled by the input variables {\tt
  ED\_USE\_KANAMORI} and {\tt ED\_READ\_UMATRIX}.

@@

\begin{tabular}{c|c}\label{list2}
{\bf normal} & {\bf superc/nonsu2}\\
\hline
  \stackinset{l}{}{t}{}{}{\listingA} & \stackinset{l}{}{t}{}{}{\listingB} 
\end{tabular}
%

>>>>

The interaction part of the impurity Hamiltonian, 
 $\hat{H}^{\mathrm{int}}$ can in principle contain any 
set of two-body operators: 
\begin{equation}
\hat{H}^\mathrm{int}_\mathrm{imp} = \frac{1}{2} \sum_{ijkl} c^{\dagger}_{i}c^{\dagger}_{j}U_{ijkl}c_{l}c_{k}.\label{HintUmat}
\end{equation}
However, we typically 
adopt a generic formulation of the local multi-orbital Hubbard-Kanamori interaction~\cite{Georges2013ACMP}:
\begin{equation}
  \begin{split}
    \hat{H}^\mathrm{int}_\mathrm{imp} &=U\sum_{\a}n_{\a\uparrow}n_{\a\downarrow}+U'\sum_{\a\neq \b}n_{\a\uparrow}n_{\b\downarrow}+(U'-J)\sum_{\a<\b,\sigma}n_{\a\sigma}n_{\b\sigma}\\
    &{\phantom =}- J_X\sum_{\a\neq
      \b}d^{+}_{\a\uparrow}d_{\a\downarrow}d^{+}_{\b\downarrow}d_{\b\uparrow}+J_P\sum_{\a
      \neq
      \b}d^{+}_{\a\uparrow}d^{+}_{\a\downarrow}d_{\b\downarrow}d_{\b\uparrow}\\
\end{split}
\label{Hint}
\end{equation}
where we introduced the occupation number operator   $n_{\alpha\sigma}=d^{+}_{\alpha\sigma}d_{\alpha\sigma}$.
The first three terms represent the density-density part of the
interaction, where $U$ is the local intra-orbital Coulomb repulsion,
$U'$ the inter-orbital one and $J$ the Hund's coupling~\cite{Georges2013ACMP}.  
The last two terms are, respectively, the spin-exchange ($J_X$) and
the pair-hopping ($J_P$).
In the three-orbital case $N_\a=3$ a fully symmetric $SU(3)_\mathrm{orbital}\otimes SU(2)_\mathrm{spin}\otimes
U(1)_\mathrm{charge}$ form of the interaction is obtained by setting $U'=U-2J$ and
$J_X=J_P=J$~\cite{Georges2013ACMP}. Different choices, preserving part
of the combined symmetry group, can be made for other number of orbitals~\cite{Georges2013ACMP}.

@@

\subsubsection{Electron-Phonon coupling}
Finally, we include a local electron-phonon coupling on the impurity site
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_\mathrm{ph}&=\sum_m \omega_{0m} b_m^+b_m   \\
    % 
    \hat{H}_\mathrm{e-ph} &= \sum_m g_m[\hat{O}_m - A_m](b_m+b_m^+) \\
\end{split}
\end{equation}
where $m=1,\dots,M$ indexes the number of local phonon modes, and $b_m$
($b_m^+$) are the destruction (creation) operators for a phonon with
frequency $\omega_{0m}$. Each phonon mode is coupled to a bilinear fermionic operator $\hat{O}_m$ on the impurity, while $A_m$ is a constant acting as a displacement field.  %Indeed, if the expectation value of the bosonic displacement operator $\langle b_q^\dagger + b_q\rangle$ is non-zero, a change of variables from the bosons $b^{(\dagger)}_q$ leads to new effective bosons $\tilde{b}^{(\dagger)}_q = b^{(\dagger)}_q- A_q$ for which the displacement operator has a smaller expectation value (ideally zero if we can impose $A_m = \langle \hat{O}_m\rangle$). 
For single-orbital impurities, the relevant case is $O_1 = n_d = \sum_{\sigma} d^{\dagger}_{\sigma}d_{\sigma}$, which defines the Holstein model\cite{Holstein1959APNY}. For two-orbital models we included also the two Jahn-Teller modes\cite{Capone2010ACMP}, for which $O_2 = n_1 - n_2$\cite{Scazzola2023PRB} and $O_3 = \sum_{\sigma} (d^{\dagger}_{1\sigma}d_{2\sigma}+ h.c.)$. For each phonon mode, one has to introduce a cut-off in the phonon number which has to be chosen according to the value electron-phonon coupling\cite{Capone2003PRL,Capone2006PRB} and its interplay with the electron-electron interaction\cite{Capone2004PRL_2,Sangiovanni2005PRB,Paci2005PRL,Sangiovanni2006PRL,Sangiovanni2006PRB_ph,Paci2006PRB}. The larger the number of excited phonons in the groundstate, the larger the cut-off. For practical calculations around 20 phonons are sufficient to achieve converged results except for the deep polaronic regime\cite{Capone2003PRL}.
Choosing properly the shift $A_m$ one can reduce the number of excited phonons in the groundstate and consequently the required cut-off. 
Although feasible, dealing with more than one phonon mode becomes quickly computationally very demanding, thus in the rest of the this work we will consider only one mode.

The phonon frequency is passed to the \NAME program via the variable { \tt W0\_PH } to be read from the input file. The displacement field acting on the phonon is passed through the variable {\tt A\_PH} from the input file. For every phonon mode, the $g_m$ coupling can be written as a matrix $g_{\alpha \beta}$ which can be passed in two different ways. If the matrix is diagonal ($g_{\alpha \beta} = g_\alpha \delta_{\alpha \beta}$), from the input file one can simply write in order the values $g_\alpha$ in the variable {\tt G\_PH }. For any general shape one can pass to the input file a {\tt GPHfile } variable containing the name of the file storing the $g_{\alpha \beta}$ array in Fortran style.
If {\tt GPHfile=NONE} the matrix $g_{\alpha \beta}$ is read from the variable {\tt G\_PH}, else it is read from the {\tt GPHfile}.

\subsubsection{System setup}
We consider a quantum impurity system composed by a single multi-orbital impurity, a bath discretized into a number $N_\mathrm{bath}$ of {\it elements} possibly endowed with an internal structure, and a number of available phonons. 
The total size of the system is determined by three contributions: i) the number of impurity orbitals $N_\a$, ii) the total count of electronic bath levels $N_\mathrm{b}$, and iii) by maximum  number of allowed phonons $N_\mathrm{ph}$, introducing a cut-off to the unbounded dimensions of the local phonon Hilbert space.
The number of bath degrees of freedom $N_\mathrm{b}$ is a
function of the bath topology and of $N_\mathrm{bath}$, which is determined internally by \NAME.
For instance, for a simple case of a purely electronic problem, each bath
element corresponds to an independent electronic level coupled to the
impurity, thus $N_\mathrm{b}\equiv N_\mathrm{bath}$. 
The total number of electronic levels in the quantum impurity systems is indicated with $N_\mathrm{s}$.

@@

\subsection{The Fock basis states}\label{sSecBasis}
The Fock space of the quantum impurity problem is defined as
$\FF=\FF_\mathrm{e}\otimes \FF_\mathrm{ph}$, where: $$
\FF_\mathrm{e}=\bigoplus_{n=0}^{N_\mathrm{s}}
S_-\HH_\mathrm{e}^{\otimes n}
$$ 
is the electronic Fock space built 
from the local electronic Hilbert space $\HH_\mathrm{e}$, and 
$$
\FF_\mathrm{ph}=\bigoplus_{n=0}^{N_q}S_+\HH_\mathrm{ph}^{\otimes n}
$$
is the
phonon Fock space, with local phonon Hilbert space
$\HH_\mathrm{ph}=\{\ket{0},\ket{1},\dots,\ket{N_\mathrm{ph}}\}$. 
Here ($S_-$) $S_+$ is the (anti-)symmetrization operator.  
%
The total dimension of the Fock space is
$$
D=4^{N_\mathrm{s}} (N_\mathrm{ph}+1)=D_\mathrm{e}D_\mathrm{ph},
$$ 
highlighting the exponential
growth with the number of electronic levels.

@@

The electronic part of the Fock state $\ket{\vec{n}}$ is represented
as a binary string of length $2N_\mathrm{s}$. Thus, any such state can be encoded
in a computer using a sequence of $2N_\mathrm{s}$ bits, or equivalently, as an integer $I=0,\dots 2^{2N_\mathrm{s}}-1$ such that $\ket{\vec{n}}=\ket{I}$.  
The electornic destruction and creation operators, $c_{a\sigma}$ and $c^+_{a\sigma}$ 
respectively, act on the Fock space as:
\begin{align*}
  c_{a\sigma}\ket{\vec{n}} &=
    \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!-\!1,\dots}
      &\text{if $n_{a\sigma}\!=\! 1$}\\
      0 &\text{otherwise}
    \end{cases};\\
    c^{+}_{a\sigma}\ket{\vec{n}} &=
     \begin{cases}
      (-1)^{\#_{a\sigma}}\ket{\dots,n_{a\sigma}\!+\!1,\dots}
      & \text{if $n_{a\sigma}\!=\! 0$}\\
      0 & \text{otherwise}
    \end{cases}    
\end{align*}
where $\#_{a\sigma}=\sum_{b\sigma'<a\sigma} n_{b\sigma'}$ accounts for the fermionic sign imposed by the Pauli principle.
The bosonic operators $b$ and $b^\dagger$ act on the phonon part of $\ket{p}$ of the Fock state as:
\begin{align*}
  b\ket{p} &= \sqrt{p-1}\ket{p} \\
  b^\dagger\ket{p} &= \sqrt{p}\ket{p} \\
\end{align*}
The implementation of Fock states and operators can be found in 
{\tt ED\_AUX\_FUNX}. In particular, in this module we define the bitwise action 
of fermionic creation and annihilation operators as functions 
{\tt CDG} and {\tt C}, used throughout the code. Additionally, 
we provide the function {\tt bdecomp} for reconstructing the Fock  state bit sequence from the integer representation.

\begin{lstlisting}[style=fstyle,numbers=none]
  subroutine C(pos,in,out,fsgn) !$c_\mathrm{pos}|\mathrm{in}\rangle=|\mathrm{out}\rangle$
    integer,intent(in)    :: pos  !Position to be set to 1
    integer,intent(in)    :: in   !Input index
    integer,intent(inout) :: out  !Output index
    real(8),intent(inout) :: fsgn !Fermionic sign   
    integer               :: l
    if(.not.btest(in,pos-1))stop "C error: C_i|...0_i...>"
    fsgn=1d0
    do l=1,pos-1
       if(btest(in,l-1))fsgn=-fsgn
    enddo
    out = ibclr(in,pos-1)         !Set bit at pos to 0
  end subroutine C
\end{lstlisting}

\begin{lstlisting}[style=fstyle,numbers=none]
  subroutine CDG(pos,in,out,fsgn) !$c^\dagger_\mathrm{pos}|\mathrm{in}\rangle=|\mathrm{out}\rangle$
    integer,intent(in)    :: pos  !Position to be set to 1
    integer,intent(in)    :: in   !Input index
    integer,intent(inout) :: out  !Output index
    real(8),intent(inout) :: fsgn !Fermionic sign   
    integer               :: l
    if(btest(in,pos-1))stop "C^+ error: C^+_i|...1_i...>"
    fsgn=1d0
    do l=1,pos-1
       if(btest(in,l-1))fsgn=-fsgn
    enddo
    out = ibset(in,pos-1)         !Set bit at pos to 1
  end subroutine CDG
\end{lstlisting}

@@

\subsection{Conserved quantum numbers}\label{sSecQNs}
To mitigate the exponential scaling of the Fock space dimension, it 
is essential to exploit suitable symmetries. Specifically, if the 
Hamiltonian $H$ commutes with a set of operators ${\cal \vec{Q}}=[{\cal Q}_1,{\cal Q}_2,\dots,{\cal Q}_M]$ such that 
$[H, {\cal Q}_i] = 0$, the Fock space can be decomposed into smaller, 
disjoint symmetry sectors labeled by quantum numbers $\vec{Q}$.

In the context of quantum impurity problems, two commonly used 
symmetries are: i) conservation of the total charge $N$ and ii) 
conservation of the total magnetization $S_z$. Although the total 
spin operator $S^2$ may also be conserved, its implementation is 
computationally challenging and typically provides only marginal 
gains, making it less practical in many cases.
As the symmetries applies to the electronic terms, here we discuss the organization of the electronic Fock states only. The presence of phonon is accounted for exploiting the tensor structure of the global Fock space.

In \NAME, three different symmetry configurations are available, 
controlled by the input variable {\tt ed\_mode} = {\bf normal}, 
{\bf superc}, or {\bf nonsu2}. A summary of the sector properties is provided in
Table~(\ref{TabSector}).

\begin{itemize}
\item{} The {\bf normal} case conserves both the electrons total occupation $N$
and total magnetization $S_z$, or equivalently, the total number
of electrons with spin up $N_\up$ and down $N_\dw$.
Optionally, the symmetry
can be extended to act independently to each orbital and spin
component, i.e.~$\vec{N_\s}=[N^1_\s,\dots,N^{N_\a}_\s]$. This
specific case is discussed extensively
Ref.~\cite{Amaricci2022} so we will not be further cover it here recalling that any consideration concerning the {\tt normal} case extends directly to this case as well.

@@

\item{} The {\bf nonsu2} case conserves the total
charge $N$ while allowing spin symmetry group to be not conserved.
This scenario captures effects such as local spin-orbit coupling
$\vec{L}\cdot\vec{S}$~\cite{something}, in-plane spin ordering~\cite{BellomiaKMH} or
in-plane spin-triplet exciton condensation~\cite{Amaricci2023_excitons,Blason}.  
\end{itemize}

\begin{table}%[ht]
\begin{center}
\begin{tabularx}{\linewidth}{ |X|X|X| } 
 \hline
  {\tt ed\_mode} & {\it Quantum Numbers} & {\it Sector Dimension} \\
  \hline
  {\bf normal} & $[N,S_z]\equiv[N_\up,N_\dw]$ &
                                                $\binom{N_\mathrm{s}}{N_\up}\binom{N_\mathrm{s}}{N_\dw}$
  \\
  \hline
  {\bf superc} & $S_z\equiv N_\up-N_\dw$ &  $\sum_m 2^{N_\mathrm{s}-S_z-2m}\binom{N_\mathrm{s}}{N_\mathrm{s}-S_z-2m}\binom{S_z+2m}{m}$
  \\
  \hline
  {\bf nonsu2} & $N \equiv N_\up+N_\dw$ & $\binom{2N_\mathrm{s}}{N}$ \\ 
 \hline
  \label{TabSector}
\end{tabularx}
\end{center}
\caption{
  {\bf Quantum Numbers Table}.
A table summarizing the possible values of the input variable {\tt
  ed\_mode} selecting the symmetries of the problem (first column). To each value
correspond different quantum numbers listed in the second column.
The third column reports the dimension of the electronic symmetry sector $\SS_{\vec{Q}}$ for a
given value of the quantum numbers $\vec{Q}$.}
\end{table}
From a computational perspective, constructing a symmetry
sector amounts to defining a injective map
$\MM:\SS_{\vec{Q}}\rightarrow \FF_e$ which relates the electronic states $\ket{i}$
within a given symmetry sector $\SS_{\vec{Q}}$ to the states $\ket{I_e}$ in the Fock space.
This map is typically implemented as a rank-1
integer array whose size corresponds to the dimension of the sector $D_{\SS_{\vec{Q}}}$. 
The global Fock state $\ket{I}=\ket{I_e}\ket{p}$, including the phonon part, is univocally determined by the relation: $I\to I_e + p 2^{2 N_s}$.

The {\bf normal} case deserves a special note. Since $N_\up$ and
$N_\dw$ are conserved independently, the local Hilbert
space and the electronic Fock space can be factorized as
$\HH_e=\HH_{e\up}\otimes\HH_{e\dw}$ and $\FF_\mathrm{e} = \FF_{e\up}\otimes \FF_{e\dw}$, respectively.  
Consequently, each electronic Fock state can be written as a product $\ket{\vec{n}_\up}\otimes \ket{\vec{n}_\dw}$. This factorization splits the
symmetry sector as $\SS_{\vec{Q}} = \SS_{N_\up}\otimes\SS_{N_\dw}$,
and finally the sector map can be expressed as the product $\MM = \MM_\up
\otimes \MM_\dw$.
%
The electronic part of each sector state $\ket{i}=\ket{i_\up} \otimes \ket{i_\dw}$ in this factorized basis
is labeled by two integers $[i_\up,i_\dw]$, with 
$i_\sigma=1,\dots,D_{\SS_\sigma}$ such that $i=i_\up + i_\dw
D_{\SS_\dw}$.
The maps $\MM_\sigma$ then connect these basis states to Fock states
$\ket{I_e}=\ket{I_\up}\ket{I_\dw}$, labeled by two integers
$[I_\up,I_\dw]$ as $I_e=I_\up +   I_\dw 2^{N_\mathrm{s}}$.
For a more detailed discussion on the structure of the Fock basis in this case, see Ref.~\cite{Amaricci2022}.

The presence of a symmetry reduces the electronic Hamiltonian matrix to
a block diagonal form, where each block, labeled by $\vec{Q}$, has dimension
$D_{\SS_{\vec{Q}}}$. The sector Hamiltonian matrix $H_\SS$ is represented in the
basis $\ket{i}\in\SS_{\vec{Q}}$ as a sparse matrix.
In the {\bf normal} case this block structure is particularly
symmetric due to the factorized nature of the sectors as discussed
above, see also Ref.~\cite{Amaricci2022}.
The analysis of the spectrum is then reduced to inspecting the
Hamiltonian within each symmetry sector, or to a subset thereof if a
additional constraints are present.

@@

In \NAME the implementation of symmetry sectors is managed through the
{\tt sector} object, which is defined in {\tt
  ED\_VARS\_GLOBAL}. This object contains all the relevant
information for defining the symmetry, including the sector
dimensions, quantum numbers and the map $\MM$. The constructor
(destructor) for this object is defined in the {\tt
  ED\_SECTORS} module via the function {\tt build\_sector} ({\tt
  delete\_sector}).
These functions use different algorithms depending on the nature 
of the quantum numbers $\vec{Q}$, i.e. the value of {\tt ed\_mode}. The following  code snippet summarize the various available implementations. The core idea is to loop of necessary Fock states through the integers $I_\sigma=1,\dots,2^{N_s}$ and enforce the quantum number constraint:

@@

\vspace{2mm}
\begin{tabular}{c|c|c}\label{list1}
    {\bf normal} & {\bf superc} & {\bf nonsu2}\\
    \hline
\stackinset{l}{}{t}{}{}{\listingA} &
\stackinset{l}{}{t}{}{}{\listingB} &
\stackinset{l}{}{t}{}{}{\listingC} \\
\end{tabular}
\vspace{2mm}

@@

\subsection{Interaction setup}\label{sSecIntSetup}
\NAME offers two distinct methods to setup the interaction terms $\hat{H}^\mathrm{int}$ in \equ{Himp}, which are controlled by the input variables {\tt
  ED\_USE\_KANAMORI} and {\tt ED\_READ\_UMATRIX}.

@@

\begin{tabular}{c|c}\label{list2}
{\bf normal} & {\bf superc/nonsu2}\\
\hline
  \stackinset{l}{}{t}{}{}{\listingA} & \stackinset{l}{}{t}{}{}{\listingB} 
\end{tabular}
\vspace{1mm}
